Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow

### 2012-12-16 Text/Submission/introduction.tex
Add Google affil to the submission text.
Finished my corrections.

\begin{document}
\begin{frontmatter}
\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}
\author[uofa]{A.~Kuzmin\corref{cor1}}
\cortext[cor1]{Corresponding author}
\ead{kuzmin@ualberta.ca}
\author[us]{M.~Januszewski}
\ead{michalj@gmail.com}
\author[schlum]{D.~Eskin}
\ead{deskin@slb.com}
\author[schlum]{F.~Mostowfi}
\ead{fmostowfi@slb.com}
\author[uofa]{J.J.~Derksen}
\ead{jos@ualberta.ca}
\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
116 St, Edmonton, Alberta, T6G
2V4 Canada}
\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
Canada}
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in a slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration to estimate mass transfer coefficients  for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent volume averaged concentration taken as the characteristic concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}
\begin{keyword}
Mass Transfer \sep Taylor/Bretherton bubble train flow \sep Multiphase flow \sep Lattice Boltzmann
method
\sep Binary liquid model \sep Flow in microchannels with parallel plates 
%% keywords here, in the form: keyword \sep keyword

@@

The understanding of mass transfer for the bubble train flow is not possible without knowledge of hydrodynamic patterns.
There are several works studying the hydrodynamic properties of the bubble train flow, both
experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}.
 For the flow of long bubbles between parallel plates chosen here as the study case, it is indicated that  there exists a vortex 
in the liquid slug for $Ca<0.7$, see Fig. \ref{fig:streamlines:tweaked:velocity}, and that the bubble shape is symmetric for low capillary 
numbers ($Ca<0.1$ \cite{cerro-bubble-train}) with the capillary number defined as:
 \beq
 Ca=\frac{\mu_{\mathrm{liq}} \ububble}{\gamma},
 \feq
 where $\mu_{\mathrm{liq}}$ is the liquid viscosity, $\ububble$ is the bubble velocity, and $\gamma$ is the interfacial tension.
 The fact that the bubble shape for $Ca<0.1$ can be represented as two hemicircles and two planar interfaces with the vortex existing in the liquid slug has been 
utilized for analytical estimations of mass transfer properties.

@@

Thus, before performing the mass transfer simulations an additional single phase hydrodynamic
simulation is performed. A free-surface solver was developed in order to obtain a velocity field
consistent with the advection-diffusion equation. We take results from the
multiphase simulations, extract a
bubble shape using the phase indicator $\phi\leq0$, and approximate the bubble shape by the stair-case
line with imposed free-slip boundary condition on it. The obtained bubble velocity is then imposed on the walls. This corresponds to conducting 
simulations in the reference frame moving with the bubble. Appendix \ref{appendix-free-surface} covers the simple
free-slip boundary condition implementation drastically improving velocity patterns. The system is
iterated until a
steady state is reached. Note, that these type of simulations are much faster than the original
multiphase simulations. As the output all the non-zero velocity components perpendicular to the
bubble surface are completely eliminated. We compared original multiphase simulations with
one-component free-slip simulations. All quantities such as superficial slug and liquid velocities are
within $3\%$ for the capillary number in the range $0.05\leq Ca \leq 1.0$. 
One can see in Fig. \ref{fig:free:surface} two streamline profiles for $Ca=0.097$  and $Ca=1.040$.
\begin{figure}[htb!]
\includegraphics[angle=90,width=\textwidth]{performed_ca0097_out.eps}\\
\includegraphics[angle=90,width=\textwidth]{performed_ca1040_out.eps}
\caption{The streamline patterns produced by the free-surface flow solver with simplified approximation
of the free-slip bubble surface, see Appendix \ref{appendix-free-surface}. Two completely different
velocity patterns are obtained, $Ca=0.097$ (top) and $Ca=1.040$ (bottom).
\label{fig:streamlines:tweaked:velocity}}
\end{figure}

>>>>

\begin{document}
\begin{frontmatter}
\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}
\author[uofa]{A.~Kuzmin\corref{cor1}}
\cortext[cor1]{Corresponding author, Telephone: +1(438)824-3695}
\ead{kuzmin@ualberta.ca}
\author[us,google]{M.~Januszewski}
\ead{michalj@gmail.com}
\author[schlum]{D.~Eskin}
\ead{deskin@slb.com}
\author[schlum]{F.~Mostowfi}
\ead{fmostowfi@slb.com}
\author[uofa]{J.J.~Derksen}
\ead{jos@ualberta.ca}
\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
116 St, Edmonton, Alberta, T6G
2V4 Canada}
\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
Canada}
\address[google]{Google Switzerland GmbH, Brandschenkestrasse 110, 8002 Zurich, Switzerland}
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in a slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration to estimate mass transfer coefficients  for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent volume averaged concentration taken as the characteristic concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}
\begin{keyword}
Mass Transfer \sep Taylor/Bretherton bubble train flow \sep Multiphase flow \sep Lattice Boltzmann
method
\sep Binary liquid model \sep Flow in microchannels with parallel plates 
%% keywords here, in the form: keyword \sep keyword

@@

The understanding of mass transfer for the bubble train flow is not possible without knowledge of hydrodynamic patterns.
There are several works studying the hydrodynamic properties of the bubble train flow, both
experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}.
 For the flow of long bubbles between parallel plates chosen here as the study case, it is indicated that  there exists a vortex 
in the liquid slug for $Ca<0.7$, %see Fig. \ref{fig:streamlines:tweaked:velocity}, 
and that the bubble shape is symmetric for low capillary 
numbers ($Ca<0.1$ \cite{cerro-bubble-train}) with the capillary number defined as:
 \beq
 Ca=\frac{\mu_{\mathrm{liq}} \ububble}{\gamma},
 \feq
 where $\mu_{\mathrm{liq}}$ is the liquid viscosity, $\ububble$ is the bubble velocity, and $\gamma$ is the interfacial tension.
 The fact that the bubble shape for $Ca<0.1$ can be represented as two hemicircles and two planar interfaces with the vortex existing in the liquid slug has been 
utilized for analytical estimations of mass transfer properties.

@@

Thus, before performing the mass transfer simulations an additional single phase hydrodynamic
simulation is performed. A free-surface solver was developed in order to obtain a velocity field
consistent with the advection-diffusion equation. We take results from the
multiphase simulations, extract a
bubble shape using the phase indicator $\phi\leq0$, and approximate the bubble shape by the stair-case
line with imposed free-slip boundary condition on it. The obtained bubble velocity is then imposed on the walls. This corresponds to conducting 
simulations in the reference frame moving with the bubble. Appendix \ref{appendix-free-surface} covers the simple
free-slip boundary condition implementation drastically improving velocity patterns. The system is
iterated until a
steady state is reached. Note, that these type of simulations are much faster than the original
multiphase simulations. As the output all the non-zero velocity components perpendicular to the
bubble surface are completely eliminated. We compared original multiphase simulations with
one-component free-slip simulations. All quantities such as superficial slug and liquid velocities are
within $3\%$ for the capillary number in the range $0.05\leq Ca \leq 1.0$. 
One can see in Fig. \ref{fig:streamlines:tweaked:velocity} two streamline profiles for $Ca=0.097$  and $Ca=1.040$.
\begin{figure}[htb!]
\includegraphics[angle=90,width=\textwidth]{performed_ca0097_out.eps}\\
\includegraphics[angle=90,width=\textwidth]{performed_ca1040_out.eps}
\caption{The streamline patterns produced by the free-surface flow solver with simplified approximation
of the free-slip bubble surface, see Appendix \ref{appendix-free-surface}. Two completely different
velocity patterns are obtained, $Ca=0.097$ (top) and $Ca=1.040$ (bottom).
\label{fig:streamlines:tweaked:velocity}}
\end{figure}
